Reinforcement Learning Report

### 2022-08-07 report/rl_qf_report.tex
update PPO algorithm

\section{RL with Actor-Critic}
\subsection{The Algorithm}
In this project, we consider the episodic learning procedure, where each 
episode has finite time steps, \(T\), to be specified as a constant, and we 
will also reset the environment at the start of each episode. The algorithm 
below is to use RL, with actor-critic method, and with the help of what we 
have built in previous sections, to solve the problem:
\begin{algorithm}[H]
\caption{Actor-Critic for our financial problem}
\begin{algorithmic}[1]
\State Initialize ANN policy parameterization 
\(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
\State Initialize ANN state-value function parameterization
\(\hat{v}(s,\bm{w})\) with any \(\bm{w}\).
\State Step sizes \(\alpha_{\bm{\theta}}>0\), \(\alpha_{\bm{w}}>0\)
\Loop \hspace{0.3mm} (for each episode)
    \State Reset the environment
    \State Generate an \(x\) (first customer of the episode)
    \State Initialize \(S\) (first state of the episode)
    \Loop \hspace{0.3mm} (for each time step \(t\), \(T\) in total)
        \State \(c\sim\pi\left(\cdot\middle|S, \bm{\theta}\right)\)
        \State Take action \(c\), observe \(S'\), \(R\) from environment
        \State \(\delta \gets R + \gamma\hat{v}(S',\bm{w})-\hat{v}(S,\bm{w})\)
        \State \(\bm{w} \gets \bm{w} + 
        \alpha_{\bm{w}} \delta\nabla\hat{v}(S,\bm{w})\)
        \State \(\bm{\theta} \gets \bm{\theta} 
        + \alpha_{\bm{\theta}}\gamma^{t}\delta
        \nabla\ln{\pi\left(A\middle|S, \bm{\theta}\right)}\)
        \State \(S \gets S'\)
    \EndLoop
\EndLoop
\end{algorithmic}
\end{algorithm}
\noindent The following graph illustrates the algorithm:
\vspace{5mm}
\begin{center}
    \begin{tikzpicture}[node distance=2cm]
        \node (agent) [process, minimum width=6cm, minimum height=3cm,
            label={[anchor=north east,inner sep=7pt]north east:Agent}]{};
        \node (pi) [process, left of=agent, xshift=0.5cm, yshift=-0.3cm] 
            {\(\pi\left(c\middle|s,\bm{\theta}\right)\)};
        \node (v) [process, right of=pi, xshift=1cm] 
            {\(\hat{v}\left(s,\bm{w}\right)\)};
        \node (env) [process, below of=pi, yshift=-2.5cm] {Environment};

@@

\section{Actor-Critic, Reinforce, and PPO}
To further show how Actor-Critic algorithm performs, we compare it to another
two popular algorithms: Reinforce and PPO. Below are pseudocodes for the two
algorithms.
\begin{algorithm}[H]
\caption{Reinforce for our financial problem}
\begin{algorithmic}[1]
\State Initialize ANN policy parameterization 
\(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
\State Step sizes \(\alpha_{\bm{\theta}}>0\)
\Loop \hspace{0.3mm} (for each episode)
    \State Reset the environment
    \State Generate an episode \(S_0, A_0, R_1, ..., S_{T-1}, A_{T-1}, R_T\),
    following \(\pi\left(\cdot\middle|\cdot, \bm{\theta}\right)\)
    \Loop \hspace{0.3mm} (for each time step \(t\), \(T\) in total)
        \State \(G \gets \sum_{k=t+1}^{T} \gamma^{k-t-1}R_k\)
        \State \(\bm{\theta} \gets \bm{\theta} + \alpha_{\bm{\theta}}\gamma^t
        G\nabla{\ln{\pi\left(A_t\middle|S_t, \bm{\theta}\right)}}\)
    \EndLoop
\EndLoop
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
\caption{PPO for our financial problem}
\begin{algorithmic}[1]
\State Initialize ANN policy parameterization 
\(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
\State Initialize ANN state-value function parameterization
\(\hat{v}(s,\bm{w})\) with any \(\bm{w}\).
\State Step sizes \(\alpha_{\bm{\theta}}>0\), \(\alpha_{\bm{w}}>0\)
\For{\(k=0, 1, 2, ...\)}
    \State Collect set of trajectories \(\mathcal{D}_k=\{\tau_i\}\), 
    following \(\pi\left(c\middle|s, \bm{\theta}\right)\)
    \State Compute discounted returns \(G_k\), the same way as in Reinforce
    algorithm
\EndFor
\end{algorithmic}
\end{algorithm}

>>>>

\section{RL with Actor-Critic}
\subsection{The Algorithm}
In this project, we consider the episodic learning procedure, where each 
episode has finite time steps, \(T\), to be specified as a constant, and we 
will also reset the environment at the start of each episode. The algorithm 
below is to use RL, with actor-critic method, and with the help of what we 
have built in previous sections, to solve the problem:
\begin{algorithm}[H]
\caption{Actor-Critic for our financial problem}
\begin{algorithmic}[1]
\State Initialize ANN policy parameterization 
\(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
\State Initialize ANN state-value function parameterization
\(\hat{v}(s,\bm{w})\) with any \(\bm{w}\).
\State Step sizes \(\alpha_{\bm{\theta}}>0\), \(\alpha_{\bm{w}}>0\)
\Loop \hspace{0.3mm} (for each episode)
    \State Reset the environment
    \State Generate an \(x\) (first customer of the episode)
    \State Initialize \(S\) (first state of the episode)
    \For{\(t=0, 1, ..., T\)}
        \State \(c\sim\pi\left(\cdot\middle|S, \bm{\theta}\right)\)
        \State Take action \(c\), observe \(S'\), \(R\) from environment
        \State \(\delta \gets R + \gamma\hat{v}(S',\bm{w})-\hat{v}(S,\bm{w})\)
        \State \(\bm{w} \gets \bm{w} + 
        \alpha_{\bm{w}} \delta\nabla\hat{v}(S,\bm{w})\)
        \State \(\bm{\theta} \gets \bm{\theta} 
        + \alpha_{\bm{\theta}}\gamma^{t}\delta
        \nabla\ln{\pi\left(A\middle|S, \bm{\theta}\right)}\)
        \State \(S \gets S'\)
    \EndFor
\EndLoop
\end{algorithmic}
\end{algorithm}
\noindent The following graph illustrates the algorithm:
\vspace{5mm}
\begin{center}
    \begin{tikzpicture}[node distance=2cm]
        \node (agent) [process, minimum width=6cm, minimum height=3cm,
            label={[anchor=north east,inner sep=7pt]north east:Agent}]{};
        \node (pi) [process, left of=agent, xshift=0.5cm, yshift=-0.3cm] 
            {\(\pi\left(c\middle|s,\bm{\theta}\right)\)};
        \node (v) [process, right of=pi, xshift=1cm] 
            {\(\hat{v}\left(s,\bm{w}\right)\)};
        \node (env) [process, below of=pi, yshift=-2.5cm] {Environment};

@@

\section{Actor-Critic, Reinforce, and PPO}
To further show how Actor-Critic algorithm performs, we compare it to another
two popular algorithms: Reinforce and PPO. Below are pseudocodes for the two
algorithms:
\begin{algorithm}[H]
\caption{Reinforce for our financial problem}
\begin{algorithmic}[1]
\State Initialize ANN policy parameterization 
\(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
\State Step sizes \(\alpha_{\bm{\theta}}>0\)
\Loop \hspace{0.3mm} (for each episode)
    \State Reset the environment
    \State Generate an episode \(S_0, C_0, R_1, ..., S_{T-1}, C_{T-1}, R_T\),
    following \(\pi\left(\cdot\middle|\cdot, \bm{\theta}\right)\)
    \For{\(t=0, 1, ..., T\)}
        \State \(G \gets \sum_{k=t+1}^{T} \gamma^{k-t-1}R_k\)
        \State \(\bm{\theta} \gets \bm{\theta} + \alpha_{\bm{\theta}}\gamma^t
        G\nabla{\ln{\pi\left(C_t\middle|S_t, \bm{\theta}\right)}}\)
    \EndFor
\EndLoop
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
\caption{PPO for our financial problem}
\begin{algorithmic}[1]
\State Initialize ANN policy parameterization 
\(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
\State Initialize ANN state-value function parameterization
\(\hat{v}(s,\bm{w})\) with any \(\bm{w}\).
\State Step sizes \(\alpha_{\bm{\theta}}>0\), \(\alpha_{\bm{w}}>0\)
\For{iteration \(k=0, 1, 2, ...\)}
    \State Collect set of trajectories \(\mathcal{D}_k=\{\tau_i\}\), 
    following \(\pi\left(\cdot\middle|\cdot, \bm{\theta}_k\right)\), where 
    \(\tau = S_0, C_0, R_1, ..., S_{T-1}, C_{T-1}, R_T\)
    \For{\(t=0, 1, ..., T\) in each trajectory \(\tau\) in \(\mathcal{D}_k\)}
        \State \(G_t \gets \sum_{k=t+1}^{T} \gamma^{k-t-1}R_k\)
        \State \(A_t \gets G_t - \hat{v}(S_t)\)
    \EndFor
    \Loop \hspace{0.3mm} (update multiple times)
    \State \(\bm{\theta} \gets \underset{\bm{\theta}}{\text{arg\:max}}
    \frac{1}{|\mathcal{D}_k|T}\sum_{\tau\in\mathcal{D}_k}^{} \sum_{t=0}^{T} 
    \min{\left(
            \frac{\pi\left(c_t\middle|S_t, \bm{\theta}\right)}
            {\pi\left(c_t\middle|S_t, \bm{\theta}_k\right)}A_t, g(\epsilon, A_t)
    \right)}\)
    \State \(\bm{w} \gets \underset{\bm{w}}{\text{arg\:min}}
    \frac{1}{|\mathcal{D}_k|T}\sum_{\tau\in\mathcal{D}_k}^{} \sum_{t=0}^{T} 
    \Bigl(\hat{v}(S_t, \bm{w})-G_t\Bigr)^2\)
    \EndLoop
\EndFor
\end{algorithmic}
\end{algorithm}
